---
title: "Forecasting the age structure of the scientific workforce in Australia"
author: Rob J Hyndman & Kelly Nguyen
date: 2 December 2025
toc: false
format:
  presentation-beamer:
    pdf-engine: pdflatex
    knitr:
      opts_chunk:
        dev: "CairoPDF"
    template-partials:
      - before-title.tex
    include-in-header: header.tex
    keep-tex: true
    fig-width: 9
    fig-height: 4.5
highlight-style: tango
execute:
  echo: false
  message: false
  warning: false
  cache: false
abstract: |
  "I will demonstrate how to forecast the age structure of a workforce using demographic growth-balance equations that incorporate functional data components for graduate influx, international mobility, retirement, and mortality. This model adapts the stochastic population framework proposed by Hyndman and Booth (IJF, 2008), substituting the birth component at age 0 with a graduation process at any age, and integrating retirement dynamics alongside a standard mortality process. It accommodates variable graduate influx, international mobility, and aging, through functional data components.
  The model is illustrated by forecasting the future age structure of scientific disciplines in Australia, utilising a variety of data sources. Data limitations are addressed through cohort interpolation and constrained penalised regression splines in order to obtain a workable data set. A combination of local and global ARIMA models is employed to address the temporal dynamics in the time series components.
  By forecasting age structures in disciplines such as Physics and Astronomy, Mathematical Sciences, Chemical Sciences, Earth Sciences, and Biological Sciences, the model informs strategic workforce planning and policy development.
  This forecasting framework not only aids in strategic planning and policy formulation within the scientific community, but also offers a blueprint for other professional sectors seeking to understand workforce dynamics amid demographic shifts."
---

```{r setup, include=FALSE}
#| eval: true
library(targets)
source("setup.R")
```

# Labour force model
\vspace*{-0.2cm}
\begin{alertblock}{}
\centerline{$P_{i,x+1,t+1} = P_{i,x,t} - D_{i,x,t} - R_{i,x,t} + G_{i,x,t} + E_{i,x,t}$}
\end{alertblock}\vspace*{-0.2cm}
\begin{block}{}
\centerline{
  $i = $ Discipline \quad
  $x = $ Age        \quad
  $t = $ Year
}
\end{block}\vspace*{-1.05cm}
\begin{align*}
P_{i,x,t} &= \text{number of equivalent full-time workers} \\
D_{i,x,t} &= \text{number of deaths} \sim \text{Binomial}(P_{i,x,t}, q_{x,t})\\
R_{i,x,t} &= \text{number of retirements}  \sim \text{Binomial}(P_{i,x,t} - D_{i,x,t}, r_{x})\\
G_{i,x,t} &= g_xG_{i,t} = \text{number of graduates who work in discipline $i$} \\
E_{i,x,t} &= \text{other changes (career changes, migration, etc.)} 
\end{align*}\pause\vspace*{-0.65cm}

\begin{block}{}
Simulate future sample paths of $P_{i,x,t}$ by simulating future\\
$q_{x,t}$, $D_{i,x,t}$, $R_{i,x,t}$, $G_{i,t}$ and $E_{i,x,t}$
\end{block}
\vspace*{10cm}

# Working population: $P_{i,x,t}$
\showmodel

```{r}
#| label: labour-force2
tar_read(fig19)
```

# Working population: $P_{i,x,t}$
\showmodel

```{r}
#| label: labour-force3
tar_read(fig20)
```

# Death probability: $q_{x,t}$
\showmodel
\begin{textblock}{5.1}(10.7,.7)\fontsize{12.5}{12.5}\sf
\begin{alertblock}{}
$D_{i,x,t} \sim \text{Binomial}(P_{i,x,t}, q_{x,t})$
\end{alertblock}
\end{textblock}

```{r}
#| label: death_prob
tar_read(fig_mxt)
```

# Death probability: $q_{x,t}$
\showmodel
\begin{textblock}{5.1}(10.7,.7)\fontsize{12.5}{12.5}\sf
\begin{alertblock}{}
$D_{i,x,t} \sim \text{Binomial}(P_{i,x,t}, q_{x,t})$
\end{alertblock}
\end{textblock}
\vspace*{0.8cm}

```{r}
#| label: fdm_model1
#| fig-height: 4.2
tar_read(fig_model_fdm1)
```

\begin{textblock}{5.2}(0.3,6.2)
\begin{block}{}\fontsize{12}{13}\sf
\centerline{$q_{x,t} = \mu(x) + \displaystyle\sum_{k=1}^6 \beta_{k,t}\phi_k(x) + \varepsilon_t(x)$}
$\beta_{k,t} \sim$ ARIMA
\end{block}
\end{textblock}
\begin{textblock}{1}(0.4,2.0)\fontsize{11}{13}\sf$\mu(x)$\end{textblock}
\begin{textblock}{1}(5.45,2.0)\fontsize{11}{13}\sf$\phi_1(x)$\end{textblock}
\begin{textblock}{1}(10.3,2.0)\fontsize{11}{13}\sf$\phi_2(x)$\end{textblock}
\begin{textblock}{1}(5.45,5.6)\fontsize{11}{13}\sf$\beta_{1,t}$\end{textblock}
\begin{textblock}{1}(10.3,5.6)\fontsize{11}{13}\sf$\beta_{2,t}$\end{textblock}

# Death probability: $q_{x,t}$
\showmodel
\begin{textblock}{5.1}(10.7,.7)\fontsize{12.5}{12.5}\sf
\begin{alertblock}{}
$D_{i,x,t} \sim \text{Binomial}(P_{i,x,t}, q_{x,t})$
\end{alertblock}
\end{textblock}
\vspace*{0.8cm}

```{r}
#| label: fdm_model2
#| fig-height: 4.2
tar_read(fig_model_fdm2)
```

\begin{textblock}{5.2}(0.3,6.2)
\begin{block}{}\fontsize{12}{13}\sf
\centerline{$q_{x,t} = \mu(x) + \displaystyle\sum_{k=1}^6 \beta_{k,t}\phi_k(x) + \varepsilon_t(x)$}
$\beta_{k,t} \sim$ ARIMA
\end{block}
\end{textblock}
\begin{textblock}{1}(0.4,2.0)\fontsize{11}{13}\sf$\mu(x)$\end{textblock}
\begin{textblock}{1}(5.45,2.0)\fontsize{11}{13}\sf$\phi_3(x)$\end{textblock}
\begin{textblock}{1}(10.3,2.0)\fontsize{11}{13}\sf$\phi_4(x)$\end{textblock}
\begin{textblock}{1}(5.45,5.6)\fontsize{11}{13}\sf$\beta_{3,t}$\end{textblock}
\begin{textblock}{1}(10.3,5.6)\fontsize{11}{13}\sf$\beta_{4,t}$\end{textblock}

# Death probability: $q_{x,t}$
\showmodel
\begin{textblock}{5.1}(10.7,.7)\fontsize{12.5}{12.5}\sf
\begin{alertblock}{}
$D_{i,x,t} \sim \text{Binomial}(P_{i,x,t}, q_{x,t})$
\end{alertblock}
\end{textblock}
\vspace*{0.8cm}

```{r}
#| label: fdm_model3
#| fig-height: 4.2
tar_read(fig_model_fdm3)
```

\begin{textblock}{5.2}(0.3,6.2)
\begin{block}{}\fontsize{12}{13}\sf
\centerline{$q_{x,t} = \mu(x) + \displaystyle\sum_{k=1}^6 \beta_{k,t}\phi_k(x) + \varepsilon_t(x)$}
$\beta_{k,t} \sim$ ARIMA
\end{block}
\end{textblock}
\begin{textblock}{1}(0.4,2.0)\fontsize{11}{13}\sf$\mu(x)$\end{textblock}
\begin{textblock}{1}(5.45,2.0)\fontsize{11}{13}\sf$\phi_5(x)$\end{textblock}
\begin{textblock}{1}(10.3,2.0)\fontsize{11}{13}\sf$\phi_6(x)$\end{textblock}
\begin{textblock}{1}(5.45,5.6)\fontsize{11}{13}\sf$\beta_{5,t}$\end{textblock}
\begin{textblock}{1}(10.3,5.6)\fontsize{11}{13}\sf$\beta_{6,t}$\end{textblock}

# Death probability: $q_{x,t}$
\showmodel
\begin{textblock}{5.1}(10.7,.7)\fontsize{12.5}{12.5}\sf
\begin{alertblock}{}
$D_{i,x,t} \sim \text{Binomial}(P_{i,x,t}, q_{x,t})$
\end{alertblock}
\end{textblock}

```{r}
#| label: death_prob3
tar_read(fig_mxt) +
  ggplot2::guides(color = "none") +
  ggplot2::labs(title = "Probability of death for Australians")
```

# Death probability: $q_{x,t}$
\showmodel
\begin{textblock}{5.1}(10.7,.7)\fontsize{12.5}{12.5}\sf
\begin{alertblock}{}
$D_{i,x,t} \sim \text{Binomial}(P_{i,x,t}, q_{x,t})$
\end{alertblock}
\end{textblock}

```{r}
#| label: future_mxt0
tar_read(fig_model_mxt0)
```

# Death probability: $q_{x,t}$
\showmodel
\begin{textblock}{5.1}(10.7,.7)\fontsize{12.5}{12.5}\sf
\begin{alertblock}{}
$D_{i,x,t} \sim \text{Binomial}(P_{i,x,t}, q_{x,t})$
\end{alertblock}
\end{textblock}

```{r}
#| label: future_mxt1
tar_read(fig_model_mxt1)
```

# Death probability: $q_{x,t}$
\showmodel
\begin{textblock}{5.1}(10.7,.7)\fontsize{12.5}{12.5}\sf
\begin{alertblock}{}
$D_{i,x,t} \sim \text{Binomial}(P_{i,x,t}, q_{x,t})$
\end{alertblock}
\end{textblock}

```{r}
#| label: future_mxt2
tar_read(fig_model_mxt2)
```

# Retirement rates: $r_{x}$
\showmodel
\begin{textblock}{6.1}(9.7,.7)\fontsize{12.5}{12.5}\sf
\begin{alertblock}{}
$R_{i,x,t} \sim \text{Binomial}(P_{i,x,t} - D_{i,x,t}, r_{x})$
\end{alertblock}
\end{textblock}

```{r}
#| label: retirement
tar_read(fig_r)
```

# Retirement rates: $r_{x}$
\showmodel
\begin{textblock}{6.1}(9.7,.7)\fontsize{12.5}{12.5}\sf
\begin{alertblock}{}
$R_{i,x,t} \sim \text{Binomial}(P_{i,x,t} - D_{i,x,t}, r_{x})$
\end{alertblock}
\end{textblock}

```{r}
#| label: retirementx1
tar_read(fig_rx1)
```

# Retirement rates: $r_{x}$
\showmodel
\begin{textblock}{6.1}(9.7,.7)\fontsize{12.5}{12.5}\sf
\begin{alertblock}{}
$R_{i,x,t} \sim \text{Binomial}(P_{i,x,t} - D_{i,x,t}, r_{x})$
\end{alertblock}
\end{textblock}

```{r}
#| label: retirementx
tar_read(fig_rx2)
```

# Graduate completions: $g_{x}$
\showmodel

```{r}
#| label: graduates
tar_read(fig_completions)
```

# Graduate completions: $g_{x}$
\showmodel

```{r}
#| label: graduates_ave
tar_read(fig_ave_completions)
```

# Graduate completions: $G_{i,t}$
\showmodel

```{r}
#| label: course_leavers
tar_read(fig21)
```

\only<2>{\begin{textblock}{4}(11,6.5)
\begin{block}{}
Global Model:\\ ARIMA(0,1,1) + drift
\end{block}
\end{textblock}}

# Graduate completions: $G_{i,t}$
\showmodel

```{r}
#| label: graduates_forecast
tar_read(fig_grad_sim)
```

\begin{textblock}{4}(11,6.5)
\begin{block}{}
Global Model:\\
ARIMA(0,1,1) + drift
\end{block}
\end{textblock}

# Remainder: $E_{x,t}$
\showmodel
\begin{alertblock}{}
\centerline{$E_{i,x,t} = P_{i,x+1,t+1} - P_{i,x,t} + D_{i,x,t} + R_{i,x,t} - g_{x}G_{i,t}$}
\end{alertblock}
\vspace*{10cm}

# Remainder: $E_{x,t}$
\showmodel
\begin{alertblock}{}
\centerline{$E_{i,x,t} = P_{i,x+1,t+1} - P_{i,x,t} + D_{i,x,t} + R_{i,x,t} - g_{x}G_{i,t}$}
\end{alertblock}

```{r}
#| label: error_component
#| fig-height: 4.1
tar_read(fig22)
```

\begin{textblock}{4.3}(11,6.5)\fontsize{12}{12}\sf
\begin{block}{}
Simulate using functional time series models, as for $q_{x,t}$.
\end{block}
\end{textblock}

# Population: $P_{i,x,t}$
\showmodel

```{r}
#| label: labour-force-disciplines2
tar_load(fig_Pxt_discipline)
fig_Pxt_discipline +
  patchwork::plot_annotation(title = "Working population by discipline")
```

# Population: $P_{i,x,t}$
\showmodel

```{r}
#| label: labour-force-disciplines0
tar_load(fig_Pxt_discipline0)
fig_Pxt_discipline0 +
  patchwork::plot_annotation(title = "Working population by discipline")
```

```{r}
#| label: labour-force-future-disciplines
#| output: asis
tar_load(fig_Pxt_future_discipline_talk)
for (i in seq_along(fig_Pxt_future_discipline_talk)) {
  cat("\n\n# Forecasts by discipline")
  cat("\n\\showmodel\n\n")
  print(fig_Pxt_future_discipline_talk[[i]])
  cat("\n")
}
```

# More information

\begin{textblock}{10.2}(2.9,3.6)
\begin{alertblock}{}
\fontsize{24}{28}\sf\bfseries\centering
\href{https://robjhyndman.com/asc2025}{robjhyndman.com/asc2025}
\end{alertblock}
\end{textblock}

