---
title: "Forecasting the age structure of the scientific workforce in Australia"
author: Rob J Hyndman & Kelly Nguyen
date: 30 June 2025
toc: false
format:
  presentation-beamer:
    pdf-engine: pdflatex
    knitr:
      opts_chunk:
        dev: "CairoPDF"
    template-partials:
      - before-title.tex
    include-in-header: header.tex
    keep-tex: true
    fig-width: 9
    fig-height: 4.5
highlight-style: tango
execute:
  echo: false
  message: false
  warning: false
  cache: false
abstract: |
  "I will demonstrate how to forecast the age structure of a workforce using demographic growth-balance equations that incorporate functional data components for graduate influx, international mobility, retirement, and mortality. This model adapts the stochastic population framework proposed by Hyndman and Booth (IJF, 2008), substituting the birth component at age 0 with a graduation process at any age, and integrating retirement dynamics alongside a standard mortality process. It accommodates variable graduate influx, international mobility, and aging, through functional data components.
  The model is illustrated by forecasting the future age structure of scientific disciplines in Australia, utilising a variety of data sources. Data limitations are addressed through cohort interpolation and constrained penalised regression splines in order to obtain a workable data set. A combination of local and global ARIMA models is employed to address the temporal dynamics in the time series components.
  By forecasting age structures in disciplines such as Physics and Astronomy, Mathematical Sciences, Chemical Sciences, Earth Sciences, and Biological Sciences, the model informs strategic workforce planning and policy development.
  This forecasting framework not only aids in strategic planning and policy formulation within the scientific community, but also offers a blueprint for other professional sectors seeking to understand workforce dynamics amid demographic shifts."
---

```{r setup, include=FALSE}
#| eval: true
library(targets)
source("setup.R")
```

# Ideal labour force model

\begin{alertblock}{}
\centerline{$P_{x+1,t+1} = P_{x,t} - D_{x,t} - R_{x,t} + G_{x,t} - C_{x,t} + N_{x,t}$}
\end{alertblock}
\begin{textblock}{1.7}(13.3,2.1)
\begin{block}{}
$x = $ Age \\
$t = $ Year
\end{block}
\end{textblock}

\begin{align*}
P_{x,t} &= \text{number of equivalent full-time workers} \\
D_{x,t} &= \text{number of deaths} \\
R_{x,t} &= \text{number of retirements} \\
G_{x,t} &= \text{number of graduates who work in science} \\
C_{x,t} &= \text{net number of people who have a career change} \\
N_{x,t} &= \text{net number of migrants} \\
\end{align*}

\vspace*{10cm}

# Pragmatic labour force model

\begin{alertblock}{}
\centerline{$P_{x+1,t+1} = P_{x,t}(1 - q_{x,t} - r_{x}) + g_{x} G_{t} + E_{x,t}$}
\end{alertblock}
\begin{textblock}{1.7}(13.3,2.1)
\begin{block}{}
$x = $ Age \\
$t = $ Year
\end{block}
\end{textblock}

\begin{align*}
P_{x,t} &= \text{number of equivalent full-time workers} \\
q_{x,t} &= \text{probability of death} \\
r_{x} &= \text{probability of retirement} \\
g_{x} &= \text{proportion of graduates by age} \\
G_{t} &= \text{total number of graduates in science} \\
E_{x,t} &= \text{remainder} \\
\end{align*}

\vspace*{10cm}

# Working population: $P_{x,t}$
\showmodel

```{r}
#| label: labour-force2
tar_read(fig_Pxt_census)
```

# Working population: $P_{x,t}$
\showmodel

```{r}
#| label: labour-force3
tar_read(fig_Pxt)
```

# Death rates: $q_{x,t}$
\showmodel

```{r}
#| label: death_prob
tar_read(fig_mxt)
```

# Retirement rates: $r_{x}$
\showmodel

```{r}
#| label: retirement
tar_read(fig_r)
```

# Retirement rates: $r_{x}$
\showmodel

```{r}
#| label: retirementx
tar_read(fig_rx2)
```

# Graduate completions: $G_{t}$
\showmodel

```{r}
#| label: course_leavers
tar_read(fig21b)
```

# Graduate completions: $g_{x}$
\showmodel

```{r}
#| label: graduates
tar_read(fig_completions)
```

# Graduate completions: $g_{x}$
\showmodel

```{r}
#| label: graduates_ave
tar_read(fig_ave_completions)
```

# Remainder: $E_{x,t}$
\showmodel
\begin{alertblock}{}
\centerline{$E_{x,t} = P_{x+1,t+1} - P_{x,t}(1 - q_{x,t} - r_{x}) - g_{x}G_{t}$}
\end{alertblock}
\vspace*{10cm}

# Remainder: $E_{x,t}$
\showmodel
\begin{alertblock}{}
\centerline{$E_{x,t} = P_{x+1,t+1} - P_{x,t}(1 - q_{x,t} - r_{x}) - g_{x}G_{t}$}
\end{alertblock}

```{r}
#| label: error_component
#| fig-height: 4.1
tar_read(fig_error)
```

# Forecasting models
\showmodel

\begin{center}
\begin{tabular}{rp{11cm}}
$G_{t}$   & ARIMA model of total graduates by year \\
$q_{x,t}$ & functional time series model \\
$E_{x,t}$ & functional time series model
\end{tabular}
\end{center}\pause\vspace*{1cm}

* Future sample paths of all components simulated to obtain probabilistic forecasts of $P_{x,t}$

# Forecasting models: $G_{t}$
\showmodel

```{r}
#| label: graduates_forecast
tar_read(fig_future_grads)
if (
  fabletools::model_sum(tar_read(arima)[[1]][[1]]) != "ARIMA(0,1,1) w/ drift"
) {
  stop("Wrong model")
}
```

\begin{textblock}{4}(11,7.2)
\begin{block}{}
Model: ARIMA(0,1,1) + drift
\end{block}
\end{textblock}

# Forecasting models: $q_{x,t}$
\showmodel

```{r}
#| label: death_prob2
tar_read(fig_mxt)
```

# Forecasting models: $q_{x,t}$
\showmodel
\vspace*{0.3cm}

```{r}
#| label: fdm_model1
tar_read(fig_model_fdm1)
```

\begin{textblock}{5.2}(0.3,6.2)
\begin{block}{}\fontsize{12}{13}\sf
\centerline{$q_{x,t} = \mu(x) + \displaystyle\sum_{k=1}^6 \beta_{k,t}\phi_k(x) + \varepsilon_t(x)$}
$\beta_{k,t} \sim$ ARIMA
\end{block}
\end{textblock}
\begin{textblock}{1}(0.4,1.5)\fontsize{11}{13}\sf$\mu(x)$\end{textblock}
\begin{textblock}{1}(5.45,1.5)\fontsize{11}{13}\sf$\phi_1(x)$\end{textblock}
\begin{textblock}{1}(10.3,1.5)\fontsize{11}{13}\sf$\phi_2(x)$\end{textblock}
\begin{textblock}{1}(5.45,5.2)\fontsize{11}{13}\sf$\beta_{1,t}$\end{textblock}
\begin{textblock}{1}(10.3,5.2)\fontsize{11}{13}\sf$\beta_{2,t}$\end{textblock}

# Forecasting models: $q_{x,t}$
\showmodel
\vspace*{0.3cm}

```{r}
#| label: fdm_model2
tar_read(fig_model_fdm2)
```

\begin{textblock}{5.2}(0.3,6.2)
\begin{block}{}\fontsize{12}{13}\sf
\centerline{$q_{x,t} = \mu(x) + \displaystyle\sum_{k=1}^6 \beta_{k,t}\phi_k(x) + \varepsilon_t(x)$}
$\beta_{k,t} \sim$ ARIMA
\end{block}
\end{textblock}
\begin{textblock}{1}(0.4,1.5)\fontsize{11}{13}\sf$\mu(x)$\end{textblock}
\begin{textblock}{1}(5.45,1.5)\fontsize{11}{13}\sf$\phi_3(x)$\end{textblock}
\begin{textblock}{1}(10.3,1.5)\fontsize{11}{13}\sf$\phi_4(x)$\end{textblock}
\begin{textblock}{1}(5.45,5.2)\fontsize{11}{13}\sf$\beta_{3,t}$\end{textblock}
\begin{textblock}{1}(10.3,5.2)\fontsize{11}{13}\sf$\beta_{4,t}$\end{textblock}

# Forecasting models: $q_{x,t}$
\showmodel
\vspace*{0.3cm}

```{r}
#| label: fdm_model3
tar_read(fig_model_fdm3)
```

\begin{textblock}{5.2}(0.3,6.2)
\begin{block}{}\fontsize{12}{13}\sf
\centerline{$q_{x,t} = \mu(x) + \displaystyle\sum_{k=1}^6 \beta_{k,t}\phi_k(x) + \varepsilon_t(x)$}
$\beta_{k,t} \sim$ ARIMA
\end{block}
\end{textblock}
\begin{textblock}{1}(0.4,1.5)\fontsize{11}{13}\sf$\mu(x)$\end{textblock}
\begin{textblock}{1}(5.45,1.5)\fontsize{11}{13}\sf$\phi_5(x)$\end{textblock}
\begin{textblock}{1}(10.3,1.5)\fontsize{11}{13}\sf$\phi_6(x)$\end{textblock}
\begin{textblock}{1}(5.45,5.2)\fontsize{11}{13}\sf$\beta_{5,t}$\end{textblock}
\begin{textblock}{1}(10.3,5.2)\fontsize{11}{13}\sf$\beta_{6,t}$\end{textblock}

# Forecasting models: $q_{x,t}$
\showmodel

```{r}
#| label: death_prob3
tar_read(fig_mxt) + 
  ggplot2::guides(color = "none") + 
  ggplot2::labs(title = "Probability of death for Australians")
```

# Forecasting models: $q_{x,t}$
\showmodel

```{r}
#| label: future_mxt0
tar_read(fig_model_mxt0)
```

# Forecasting models: $q_{x,t}$
\showmodel

```{r}
#| label: future_mxt1
tar_read(fig_model_mxt1)
```

# Forecasting models: $q_{x,t}$
\showmodel

```{r}
#| label: future_mxt2
tar_read(fig_model_mxt2)
```

# Remainder: $E_{x,t}$
\showmodel

```{r}
#| label: error_component2
tar_read(fig_error)
```
# Forecasting models: $E_{x,t}$
\showmodel
\vspace*{0.3cm}

```{r}
#| label: fdm_Emodel1
tar_read(fig_Emodel_fdm1)
```

\begin{textblock}{5.2}(0.3,6.2)
\begin{block}{}\fontsize{12}{13}\sf
\centerline{$E_{x,t} = \mu(x) + \displaystyle\sum_{k=1}^6 \beta_{k,t}\phi_k(x) + \varepsilon_t(x)$}
$\beta_{k,t} \sim$ ARIMA
\end{block}
\end{textblock}
\begin{textblock}{1}(0.4,1.5)\fontsize{11}{13}\sf$\mu(x)$\end{textblock}
\begin{textblock}{1}(5.45,1.5)\fontsize{11}{13}\sf$\phi_1(x)$\end{textblock}
\begin{textblock}{1}(10.3,1.5)\fontsize{11}{13}\sf$\phi_2(x)$\end{textblock}
\begin{textblock}{1}(5.45,5.2)\fontsize{11}{13}\sf$\beta_{1,t}$\end{textblock}
\begin{textblock}{1}(10.3,5.2)\fontsize{11}{13}\sf$\beta_{2,t}$\end{textblock}

# Forecasting models: $E_{x,t}$
\showmodel
\vspace*{0.3cm}

```{r}
#| label: fdm_Emodel2
tar_read(fig_Emodel_fdm2)
```

\begin{textblock}{5.2}(0.3,6.2)
\begin{block}{}\fontsize{12}{13}\sf
\centerline{$E_{x,t} = \mu(x) + \displaystyle\sum_{k=1}^6 \beta_{k,t}\phi_k(x) + \varepsilon_t(x)$}
$\beta_{k,t} \sim$ ARIMA
\end{block}
\end{textblock}
\begin{textblock}{1}(0.4,1.5)\fontsize{11}{13}\sf$\mu(x)$\end{textblock}
\begin{textblock}{1}(5.45,1.5)\fontsize{11}{13}\sf$\phi_3(x)$\end{textblock}
\begin{textblock}{1}(10.3,1.5)\fontsize{11}{13}\sf$\phi_4(x)$\end{textblock}
\begin{textblock}{1}(5.45,5.2)\fontsize{11}{13}\sf$\beta_{3,t}$\end{textblock}
\begin{textblock}{1}(10.3,5.2)\fontsize{11}{13}\sf$\beta_{4,t}$\end{textblock}

# Forecasting models: $E_{x,t}$
\showmodel
\vspace*{0.3cm}

```{r}
#| label: fdm_Emodel3
tar_read(fig_Emodel_fdm3)
```

\begin{textblock}{5.2}(0.3,6.2)
\begin{block}{}\fontsize{12}{13}\sf
\centerline{$E_{x,t} = \mu(x) + \displaystyle\sum_{k=1}^6 \beta_{k,t}\phi_k(x) + \varepsilon_t(x)$}
$\beta_{k,t} \sim$ ARIMA
\end{block}
\end{textblock}
\begin{textblock}{1}(0.4,1.5)\fontsize{11}{13}\sf$\mu(x)$\end{textblock}
\begin{textblock}{1}(5.45,1.5)\fontsize{11}{13}\sf$\phi_5(x)$\end{textblock}
\begin{textblock}{1}(10.3,1.5)\fontsize{11}{13}\sf$\phi_6(x)$\end{textblock}
\begin{textblock}{1}(5.45,5.2)\fontsize{11}{13}\sf$\beta_{5,t}$\end{textblock}
\begin{textblock}{1}(10.3,5.2)\fontsize{11}{13}\sf$\beta_{6,t}$\end{textblock}

# Forecasting models: $E_{x,t}$
\showmodel

```{r}
#| label: remainder2
tar_read(fig_error) + 
  ggplot2::guides(color = "none") + 
  ggplot2::labs(y = "Number of people", title = "Remainder")
```

# Forecasting models: $E_{x,t}$
\showmodel

```{r}
#| label: future_Ext0
tar_read(fig_model_Ext0)
```

# Forecasting models: $E_{x,t}$
\showmodel

```{r}
#| label: future_Ext1
tar_read(fig_model_Ext1)
```

# Forecasting models: $E_{x,t}$
\showmodel

```{r}
#| label: future_Ext2
tar_read(fig_model_Ext2)
```

# Final forecasts: $P_{x,t}$
\showmodel

```{r}
#| label: labour-force4
tar_read(fig_Pxt) + 
  ggplot2::ylim(-1,21000) + 
  ggplot2::guides(color = "none") +
  ggplot2::labs(title = "Working population")
```


```{r}
#| label: labour-force-future
#| output: asis
tar_load(fig_Pxt_future)
for(i in seq_along(fig_Pxt_future)) {
  cat("\n\n# Final forecasts: $P_{x,t}$")
  cat("\n\\showmodel\n\n")
  print(fig_Pxt_future[[i]] + ggplot2::ylim(-1,21000))
  cat("\n")
}
```

